Navigating Trends: Analyzing Ticket Sales and Redemption Patterns For Toronto Island Ferries

### 2024-09-23 Paper/Paper.qmd
filtered the dataset to analyse years from 2018-2024.

---
title: "Navigating Trends: Analyzing Ticket Sales and Redemption Patterns For Toronto Island Ferries"
author: Krishna Kumar
thanks: "Code and data are available at: https://github.com/krishnak30/Toronto-Island-Ferry-Ticket-Counts-Krishna-K.git."
date: today
date-format: long
abstract: "This report looks at the ferry services to the Toronto Islands, a significant recreational destination accessible exclusively by water. We analyze a data set of Ferry Ticket Counts from May 2015 to September 2024 in order to find patterns and trends in ferry usage and the effects of external factors. Using time series analysis and data visualization, the data shows that there exist definitive seasonal trends, with sharp declines during the pandemic period. Our findings suggest that ferry usage is highly seasonal and very vulnerable to external shocks. The paper concludes by discussing broader implications for tourism, urban and contingency planning, and public transportation management."
format: pdf
number-sections: true
bibliography: references.bib
execute:
  echo: false     # Hide code globally
  warning: false  # Suppress warnings globally
  message: false  # Suppress messages globally
references: true
---

@@

cleaned_ticket_counts_data <- read_csv(here::here("Data/Analysis Data/cleaned_ticket_counts_data.csv"))

@@

This paper analyzes the ferry ticket sales and redemption data from the Toronto Islands Ferry Service over the following time period: 4th of May, 2015 to the 20th of September, 2024. The Toronto Islands serve not only as a beloved spot for locals, but also as a popular tourist destination, attracting hundreds of thousands of visitors annually. Given the ferry's role as the primary means of transportation to the islands, understanding its usage patterns is critical for officials, policymakers, stakeholders, and business owners in order to improve service efficiency, prepare for fluctuating demand, accommodate future growth, and develop contingency plans. The significance of this research goes well beyond the ferry system, having impacts on issues of urban mobility, tourism management, and even economic recovery in a post-pandemic society.

@@

The paper is structured as follows: Section 2 **\[CROSS REFERENCE HERE\]** provides an in-depth discussion of the data, including its source, the critical aspects of the cleaning process, preparation for analysis, as well as discussion of the broader context. Section 3 **\[CROSS REFERENCE HERE\]** presents a detailed analysis of ticket sales and redemption patterns, highlighting key trends, seasonal variations, and the impact of external factors such as the COVID-19 pandemic. Finally, section 4 **\[CROSS REFERENCE HERE\]** discusses the implications of these findings, explores potential applications, and addresses the limitations of this study, offering suggestions for future research.

@@

The data is published by Parks, Forestry, and Recreation and is updated on an hourly basis, providing real-time data and updates. The data provides information regarding the ferries that takes visitors to the Toronto Island Park, which is considered a popular summer spot for tourists and locals alike. These ferries carry passengers to and from Centre Island, Hanlan's Point, and Ward's Island. The raw data features the counts for the total number of tickets redeemed and tickets sold in particular time periods. The time stamps are shows in 15 minute intervals and contain the data for redemptions and sales in the past 15 minutes. The time period for the data set ranges from the 4th of May, 2015 to the 20th of September, 2024 and contains around 229,000 rows/data entries.

@@

The data cleaning process involved breaking down the 'timestamp' column into two 'Date' and 'Time' columns for a better and more clear analysis. Along with changing names of the columns, deleting duplicate columns, and removing entries from the data that served as outliers.

@@

The summary statistics are provided in table 2 [@tbl-summary] and allow us to get an understanding of the data in a very comprehensible manner. The high volume of tickets sold and redeemed indicates a healthy and significant demand of the ferry and the closeness of the values conveys strong customer engagement and a reliable service. On average, 49.36 tickets were sold every 15 minutes during working hours, and 48.44 tickets were redeemed, indicating that most customers utilized their tickets. However, the small gap implies some no-shows or unused tickets. The maximum numbers of redemption and sales tickets (7,216 and 7,229) showcases the ferry's ability to handle high customer volumes and the closeness of these values tells us that, on busy days, the ferry usually operates at capacity.

@@

```{r}
### WHAT THE FUCK, HOW DO I MAKE THIS FIT IN
# SUMMARY TABLE

@@

After studying the summary statistics, we have an idea of the capacity and performance of the ferry, along with the relationships between the variables. Next, with the use fo figure 1 [@fig-timeseriesanalysis], we can study the patterns of ticket counts from 2015 to 2024. For figure1, we can see that there is a definitive pattern in the usage numbers when looking at each year. We see that there are spikes during the mid-year months for each year, indicating higher popularity of the Toronto Islands during those times and suggesting that the islands serve as a significant recreational destination.

@@

#| label: fig-timeseriesanalysis
#| fig-cap: "Tickets Sold and Redeemed by Date"
#| warning: false
#| echo: false
#| message: false
#| fig-align: center

@@

Figure 2 [@fig-monthanalysis], shows a monthly analysis of redemption and sales ticket counts for each year. The data in this figure has been grouped together and summed. i.e. the ticket counts for each of the 12 months in every year from 2015 to 2024 has been summed and aggregated so that we can see the monthly trends over the past 9 years.

@@

This paper ferry ticket data from Open Data Toronto, focusing on the trends of ticket sales and redemptions from the 4th of May, 2015 to the 20th of September, 2024. Through summary statistics and visualizations, we identified distinct patters in ferry usage and how external factors have influenced these trends. The time series analysis illustrated a consistent peak in ferry activity during the summer months, while the month-by-month backed up this claim and showed how seasonal variations impacted ticket counts.

>>>>

---
title: "Navigating Trends: Analyzing Ticket Sales and Redemption Patterns For Toronto Island Ferries"
author: Krishna Kumar
thanks: "Code and data are available at: https://github.com/krishnak30/Toronto-Island-Ferry-Ticket-Counts-Krishna-K.git."
date: today
date-format: long
abstract: "This report looks at the ferry services to the Toronto Islands, a significant recreational destination accessible exclusively by water. We analyze a data set of Ferry Ticket Counts from May 2018 to September 2024 in order to find patterns and trends in ferry usage and the effects of external factors. Using time series analysis and data visualization, the data shows that there exist definitive seasonal trends, with sharp declines during the pandemic period. Our findings suggest that ferry usage is highly seasonal and very vulnerable to external shocks. The paper concludes by discussing broader implications for tourism, urban and contingency planning, and public transportation management."
format: pdf
number-sections: true
bibliography: references.bib
execute:
  echo: false     # Hide code globally
  warning: false  # Suppress warnings globally
  message: false  # Suppress messages globally
references: true
fig-align: center 
---

@@

cleaned_ticket_counts_data <- read_csv(here::here("Data/Analysis_Data/cleaned_ticket_counts_data.csv"))

@@

This paper analyzes the ferry ticket sales and redemption data from the Toronto Islands Ferry Service over the following time period: 1st of January, 2018 to the 20th of September, 2024. The Toronto Islands serve not only as a beloved spot for locals, but also as a popular tourist destination, attracting hundreds of thousands of visitors annually. Given the ferry's role as the primary means of transportation to the islands, understanding its usage patterns is critical for officials, policymakers, stakeholders, and business owners in order to improve service efficiency, prepare for fluctuating demand, accommodate future growth, and develop contingency plans. The significance of this research goes well beyond the ferry system, having impacts on issues of urban mobility, tourism management, and even economic recovery in a post-pandemic society.

@@

The paper is structured as follows: Section 2 provides an in-depth discussion of the data, including its source, the critical aspects of the cleaning process, preparation for analysis, as well as discussion of the broader context. Section 3 presents a detailed analysis of ticket sales and redemption patterns, highlighting key trends, seasonal variations, and the impact of external factors such as the COVID-19 pandemic. Finally, section 4 discusses the implications of these findings, explores potential applications, and addresses the limitations of this study, offering suggestions for future research.

@@

The data is published by Parks, Forestry, and Recreation and is updated on an hourly basis, providing real-time data and updates. The data provides information regarding the ferries that takes visitors to the Toronto Island Park, which is considered a popular summer spot for tourists and locals alike. These ferries carry passengers to and from Centre Island, Hanlan's Point, and Ward's Island. The raw data features the counts for the total number of tickets redeemed and tickets sold in particular time periods. The time stamps are shows in 15 minute intervals and contain the data for redemptions and sales in the past 15 minutes. The time period for the data set ranges from the 1st of January, 2018 to the 20th of September, 2024 and contains around 229,000 rows/data entries.

@@

The data cleaning process involved breaking down the 'timestamp' column into two 'Date' and 'Time' columns for a better and more clear analysis. Along with changing names of the columns, deleting duplicate columns, and removing entries from the data that served as outliers. The data was filtered to only analyse data from 2018 to 2024, removing years 2015 - 2018. This was because the data set was too large to analyse altogether efficiently and the new filtered data set allowed for a closer analysis of the impact of COVID-19.

@@

The summary statistics are provided in table 2 [@tbl-summary] and allow us to get an understanding of the data in a very comprehensible manner. The high volume of tickets sold and redeemed indicates a healthy and significant demand of the ferry and the closeness of the values conveys strong customer engagement and a reliable service. On average, 46.92 tickets were sold every 15 minutes during working hours, and 47.37 tickets were redeemed, indicating that most customers utilized their tickets. However, the small gap implies some no-shows or unused tickets. The maximum numbers of redemption and sales tickets (7,216 and 7,229) showcases the ferry's ability to handle high customer volumes and the closeness of these values tells us that, on busy days, the ferry usually operates at capacity.

@@

```{r}
# SUMMARY TABLE

@@

After studying the summary statistics, we have an idea of the capacity and performance of the ferry, along with the relationships between the variables. Next, with the use fo figure 1 [@fig-timeseriesanalysis], we can study the patterns of ticket counts from 2018 to 2024. For figure1, we can see that there is a definitive pattern in the usage numbers when looking at each year. We see that there are spikes during the mid-year months for each year, indicating higher popularity of the Toronto Islands during those times and suggesting that the islands serve as a significant recreational destination.

@@

#| label: fig-timeseriesanalysis
#| fig-cap: "Tickets Sold and Redeemed by Date"
#| warning: false
#| echo: false
#| message: false
#| fig-align: "center"

@@

Figure 2 [@fig-monthanalysis], shows a monthly analysis of redemption and sales ticket counts for each year. The data in this figure has been grouped together and summed. i.e. the ticket counts for each of the 12 months in every year from 2018 to 2024 has been summed and aggregated so that we can see the monthly trends over the past 6 years.

@@

This paper ferry ticket data from Open Data Toronto, focusing on the trends of ticket sales and redemptions from the 1st of January, 2018 to the 20th of September, 2024. Through summary statistics and visualizations, we identified distinct patters in ferry usage and how external factors have influenced these trends. The time series analysis illustrated a consistent peak in ferry activity during the summer months, while the month-by-month backed up this claim and showed how seasonal variations impacted ticket counts.
